Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy

### 2020-07-17 chapters/ch_data_analysis.tex
Almost final draft

\subsection{Gating grid noise subtraction}
\label{sec:ggnoisesub}
Opening the gating grid is essentially done by shorting the adjacent wires in the gating grid together, and allowing them to reach equilibrium as fast as possible. Due to coupling between the gating grid, the ground plane, and other effects, the impedance of the supply cables was not perfectly matched, causing an imbalance in the current where one side discharged faster than the other side. This caused an underdamped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum. Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum after averaging over 2000 events in a given pad. 
 %The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.

@@

 where $\phi$ is the electric potential and $\rho$ is the free charge. The numerical solution is provided by the Jacobi method \cite{poisson}, provided all 6 sides have defined Dirchlet boundary conditions for the potentials. We neglect the wire plane region since the drift details around it are not significant for the space charge effects we are discussing here. The pad plane and cathode are trivial, where the side walls of the field cage are given linearly varying potentials on the surface, such that the electric field is the same strength as the real TPC with no space charge present. Once the electric potential is solved, the electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.

@@

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{SC_Relation.pdf}
\caption{Space charge relation}
\label{fig:spacechg_relation}
\end{figure}

@@

  \end{center}
  \caption{Cobo timing correction.}
  \label{fig:coboCorr}
\end{figure}

@@

The detailed software flow diagram of the embedding algorithm is shown in Fig.~\ref{fig:flow}. Starting from Geant4 and passing through the 3 MC digitization tasks described above, the output response of the MC track in the TPC is created. From here we directly embed the MC signals into experimental data by adding the MC signals directly into the experimental time bucket spectrum, pad-by-pad. As will be described in Sec.~\ref{sec:simSat}, prior to embedding the MC signals, all pads which are saturated in the experimental data are identified and flagged. MC signals are forbidden to be embedded after the (MC) saturation time bucket position identified in the procedure outlined in Section~\ref{sec:simSat}. Figure~\ref{fig:mcevent} shows the response of a \SI{200}{\mega\electronvolt\per\clight} $\pi^-$ track in the TPC. Figure~\ref{fig:dataevent} shows this track embedded into an experimental event. In the following, here I will refer to experimental data with embedded MC signals as ``embedded data" for short, the MC generated response as ``MC data", and data only containing only the experimental signals as ``experimental data". The three sets of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set. These three independent hit data sets are temporarily stored.

@@

Simulating the saturation effects is handled naturally by embedding. Once the time bucket of the saturating signal has been identified, no further signals are embedded since the pad is assumed to be dead. In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded for any time bucket. We only need to identify where the time of saturation occurs. The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}. The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation \cite{polezero}. If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,

@@

By applying this technique to the real data, we can identify whether a pad has a saturating pulse by looking for this negative undershoot. A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ ADC for more than 8 time buckets, and if the max ADC value is > $G\cdot$500, where $G$ is the gain calibration coming from low gain sections discussed in Section~\ref{sec:anodeCalib}. The max ADC condition eliminates false positives which come from  the dead pads where the gating grid subtraction is still applied, introducing large false negative peaks, but have no positive signals. After meeting these two conditions, the pad is flagged as saturated and the time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket position of the negative peak. This is because the falling edge is around 5 time buckets after the rising edge of the saturation. A separate time is also stored called the MC time bucket position, which sets the point in time in which a signal cannot be embedded any further. This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible. These two positions are also shown in Fig.~\ref{fig:pulseSatTag}.

@@

The other important cut which is made inside of the software is on the PRF distribution of each cluster discussed in Section~\ref{sec:corrTask}. We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Fig.~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data. It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}. Therefore these effects must arise from geometric effects related to the track angle, the amount of charge distributed over an anode wire, and the superposition of the PRF from neighboring anode wires which contribute to the appearance of a changing PRF.

@@

The area enclosed by the red lines represent the region of the highest efficiency over the full region in $\theta_{Lab}$.  Since we take cuts similar to these in the real data, we assume the efficiency is independent of $\phi$ over these small regions. The efficiency then can be represented as just a function of the two remaining variables. Figure~\ref{fig:132_eff} shows the efficiency of $\pi^-$ and $\pi^+$ particles in the $\tin{132}{124}$ system and Fig.~\ref{fig:108_eff} for the $\tin{108}{112}$ system as a function of $\theta_{Lab}$ and momentum $p_{Lab}$.

@@

where the parameters $p_0-p_4$ are free parameters, $\beta= p/\sqrt{p^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum. The PID lines vary slightly as a function of the emission angle of each track. The PID was subdivided into 6 pitch angles, $\theta_P = \arctan(p_y/p_z)$, and 6 yaw angles, $\theta_Y = \arctan(p_x/p_z)$, ranging from \ang{-90} to \ang{90} for both. The pion spectra of each yaw-pitch bin was fit with Eq.~\ref{eq:5par_bb} to get the PID line which best describes the data. The distribution around this mean value is not a Gaussian distribution. The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Guassian variable, defined as

@@

    \caption{$\langle dE/dx\rangle$ versus p/q for $\tin{132}{124}$ system.}
\label{fig:pidpion}
\end{figure}

Now that the pion distribution is flattened, we take bin slices along $p/q$ to determine the particle yield in a certain bin and also estimate the background contribution. The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}. Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Fig.~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles. A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.

@@

    \caption{$z_i$ projection for binned slices in p/q around the $\pi^-$.}
\label{fig:pimmom_flat}
\end{figure}

@@

    \caption{$z_i$ projection for binned slices in p/q around the $\pi^+$.}
\label{fig:pipmom_flat}
\end{figure}

>>>>

\subsection{Gating grid noise subtraction}
\label{sec:ggnoisesub}
Opening the gating grid is essentially done by shorting the adjacent wires in the gating grid together, and allowing them to reach equilibrium as fast as possible. Due to coupling between the gating grid, the ground plane, and other effects, the impedance of the supply cables was not perfectly matched, causing an imbalance in the current where one side discharged faster than the other side. This caused an under-damped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum. Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum after averaging over 2000 events in a given pad. 
 %The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.

@@

 where $\phi$ is the electric potential and $\rho$ is the free charge. The numerical solution is provided by the Jacobi method \cite{poisson}, provided all 6 sides have defined Dirichlet boundary conditions for the potentials. We neglect the wire plane region since the drift details around it are not significant for the space charge effects we are discussing here. The pad plane and cathode are trivial, where the side walls of the field cage are given linearly varying potentials on the surface, such that the electric field is the same strength as the real TPC with no space charge present. Once the electric potential is solved, the electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.

@@

\begin{figure}[!htb]
\includegraphics[width=\linewidth]{SC_Relation.pdf}
\caption{Relationship between the observable  $\Delta\mathrm{V}_\mathrm{LR}$ and the estimated charge density, and the beam intensity for each run. The blue and green line represents two separate runs at different beam intensities. The condition for  $\Delta\mathrm{V}_\mathrm{LR} = 0$ is solved for for each run. These values are then plotted against the beam intensity where finally a linear fit provides the interpolation for all beam intensities in between.}
\label{fig:spacechg_relation}
\end{figure}

@@

  \end{center}
  \caption{Cobo timing correction for three different layers. One singe pad is represented by a unique layer and row number. The $dy$ residual of the track fitting shows, in the upper panel, that there is a timing calibration issue. This offset is fixed over several runs for a given pad. Therefore we can correct using a time correction map for each pad; the bottom panels show the distribution after correction. }
  \label{fig:coboCorr}
\end{figure}

@@

The detailed software flow diagram of the embedding algorithm is shown in Figure~\ref{fig:flow}. Starting from Geant4 and passing through the 3 MC digitization tasks described above, the output response of the MC track in the TPC is created. From here we directly embed the MC signals into experimental data by adding the MC signals directly into the experimental time bucket spectrum, pad-by-pad. As will be described in Sec.~\ref{sec:simSat}, prior to embedding the MC signals, all pads which are saturated in the experimental data are identified and flagged. MC signals are forbidden to be embedded after the (MC) saturation time bucket position identified in the procedure outlined in Section~\ref{sec:simSat}. Figure~\ref{fig:mcevent} shows the response of a \SI{200}{\mega\electronvolt\per\clight} $\pi^-$ track in the TPC. Figure~\ref{fig:dataevent} shows this track embedded into an experimental event. In the following, here I will refer to experimental data with embedded MC signals as \emph{embedded data} for short, the MC generated response as \emph{MC data}, and data only containing only the experimental signals as \emph{experimental data}. The three sets of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set. These three independent hit data sets are temporarily stored.

@@

Simulating the saturation effects is handled naturally by embedding. Once the time bucket of the saturating signal has been identified, no further signals are embedded since the pad is assumed to be dead. In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded for any time bucket. We only need to identify where the time of saturation occurs. The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Figure~\ref{fig:pulseshape}. The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation \cite{polezero}. If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,

@@

By applying this technique to the real data, we can identify whether a pad has a saturating pulse by looking for this negative undershoot. A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ ADC for more than 8 time buckets, and if the max ADC value is > $G\cdot$500, where $G$ is the gain calibration coming from low gain sections discussed in Section~\ref{sec:anodeCalib}. The max ADC condition eliminates false positives which come from  the dead pads where the gating grid subtraction is still applied, introducing large false negative peaks, but have no positive signals. After meeting these two conditions, the pad is flagged as saturated and the time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket position of the negative peak. This is because the falling edge is around 5 time buckets after the rising edge of the saturation. A separate time is also stored called the MC time bucket position, which sets the point in time in which a signal cannot be embedded any further. This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible. These two positions are also shown in Figure~\ref{fig:pulseSatTag}.

@@

The other important cut which is made inside of the software is on the PRF distribution of each cluster discussed in Section~\ref{sec:corrTask}. We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Figure~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data. It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}. Therefore these effects must arise from geometric effects related to the track angle, the amount of charge distributed over an anode wire, and the superposition of the PRF from neighboring anode wires which contribute to the appearance of a changing PRF.

@@

The area enclosed by the red lines represent the region of the highest efficiency over the full region in $\theta_{Lab}$.  Since we take cuts similar to these in the real data, we assume the efficiency is independent of $\phi$ over these small regions. The efficiency then can be represented as just a function of the two remaining variables. Figure~\ref{fig:132_eff} shows the efficiency of $\pi^-$ and $\pi^+$ particles in the $\tin{132}{124}$ system and Figure~\ref{fig:108_eff} for the $\tin{108}{112}$ system as a function of $\theta_{Lab}$ and momentum $p_{Lab}$.

@@

where the parameters $p_0-p_4$ are free parameters, $\beta= p/\sqrt{p^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum. The PID lines vary slightly as a function of the emission angle of each track. The PID was subdivided into 6 pitch angles, $\theta_P = \arctan(p_y/p_z)$, and 6 yaw angles, $\theta_Y = \arctan(p_x/p_z)$, ranging from \ang{-90} to \ang{90} for both. The pion spectra of each yaw-pitch bin was fit with Eq.~\ref{eq:5par_bb} to get the PID line which best describes the data. The distribution around this mean value is not a Gaussian distribution. The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Gaussian variable, defined as

@@

    \caption{The charged pion PID plot achieved in the $\tin{132}{124}$ system. PID is achieve by plotting $\langle dE/dx\rangle$ versus particle rigidity, $p/q$. Unique hyperbolic lines show different particle species, where negatively charged particles are plotted as negative rigidity.  }
\label{fig:pidpion}
\end{figure}

Now that the pion distribution is flattened, we take bin slices along $p/q$ to determine the particle yield in a certain bin and also estimate the background contribution. The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}. Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Figure~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles. A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.

@@

    \caption{Flattened PID $z_i$ distribution for binned slices in $p/q$ around the $\pi^-$.}
\label{fig:pimmom_flat}
\end{figure}

@@

    \caption{Flattened PID $z_i$ distribution  for binned slices in p/q around the $\pi^+$.}
\label{fig:pipmom_flat}
\end{figure}

### 2020-07-17 chapters/ch_data_analysis_2.tex
Added bills edits working on appendix figures
Almost final draft

\chapter{Data Analysis II}
As Chapter~\ref{chap:data1} covered the corrections and various calibrations to the data, this Chapter will cover the various cuts taken to improve the signal in the experimental data. Some cuts which identify the incoming beam and the event on target are the minimum cut in order to perform an analysis. The cuts pertaining to the track reconstruction quality are also required to suppress as much of the background noise as possible, most of which comes from reconstruction inefficiencies which the cuts presented here address.

@@

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn132.png}
         \caption{$\tin{132}{124}$ Beam particle identification.}
         \label{fig:beampid132}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn108.png}
         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
         \label{fig:beampid108}
     \end{subfigure}
        \caption{$\tin{108}{112}$ Beam particle identification. }
        \label{fig:beampidTwo}

\end{figure}

%Figures of beam contaminants in our beam PID line. 
%Table of beam purity reference to Jon's thesis paper
The secondary beam is produced through the projectile fragmentation of the primary beam off of a \SI{3}{\milli\metre} thick, rotating Be target \cite{inflightsep}. The resulting fragments are filtered in-flight to the desired seconary beam. The in-flight separation is handled by the BigRIPS fragment separator which is shown in Fig.~\ref{fig:samuraiBeamLine}. The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$. Several sets of slits further purify the secondary beam quality by discarding particles which do not focus on the right focal planes. These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.  Each beam is tracked with the remaining part of the BigRIPS spectrometer tracking system. The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}. From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussian in this space.

@@

\begin{equation}
  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
   \label{multiGauss}
\end{equation}
where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Fig.~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{Multivariate normal distribution fit parameters for four beams.
    \label{beamParameters}}
  \end{center}
\end{table}

Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.  This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. This also occurs for the first and last time bucket in the vertical direction. While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{ccccc}
      \hline 
      System & $\mu_{\theta_\mathrm{a,proj}}$ &
      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
      $\sigma_{\theta_\mathrm{b,proj}}$ \\
      \hline\hline 
      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
      $^{108}$Sn beam angle. \label{beamAngleParameters}}
  \end{center}
\end{table}

@@

 The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame. The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC. Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}rrr@{}}\toprule
\multicolumn{3}{c}{Vertex Resolution}\\
\cmidrule{1-3}
System & Mean (cm) & Sigma (cm) \\
\midrule
$\tin{132}{124}$ & -14.79  & 1.2 \\
$\tin{124}{112}$ & -14.71  & 1.1 \\
$\tin{112}{124}$ & -14.78  & 1.2 \\
$\tin{108}{112}$ & -14.75  & 1.3 \\ 
\bottomrule
\end{tabular}
\caption{Summary of measured verticies and their resolution.}
\label{tb:vertexresol}
\end{table*}

@@

\section{Impact Parameter Selection}

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=1.]{distancetovertex.pdf}
\label{fig:poca}
\caption{Distance to vertex for tracks.}
\end{figure}

@@

Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that effect the efficiency of track reconstruction. Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}. The relevant geometries which determine these regions are the corners of the field cage. Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given. Since the window to the field cage was slightly higher, the upper angles are smaller than the lower angles. These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.  While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks. Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly. The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}. The values are very close to the simple calculation outlined previously.

@@

These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.  We could of course include a larger region where the efficiency is small, but not zero, but it is not good practice to  correct using small efficiency values. Instead we apply a reasonable $\phi$ cut in the good areas of reasonable efficiency and correct for the regions excluded from the cut. The solid angle covered by the cut is written as,

>>>>

\chapter{Data Analysis II: Cuts and final PID}
\label{chap:data2}

As Chapter~\ref{chap:data1} covered the corrections and various calibrations to the data, this Chapter will cover the various cuts taken to improve the signal in the experimental data. Some cuts which identify the incoming beam and the event on target are the minimum cut in order to perform an analysis. The cuts pertaining to the track reconstruction quality are also required to suppress as much of the background noise as possible, most of which comes from reconstruction inefficiencies which the cuts presented here address.

@@

     \centering
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn132.png}
         \caption{$\tin{132}{124}$ system.}
         \label{fig:beampid132}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.49\textwidth}
         \centering
         \includegraphics[width=\textwidth]{beam-Sn108.png}
         \caption{$\tin{108}{112}$ system.}
         \label{fig:beampid108}
     \end{subfigure}
        \caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
        \label{fig:beampidTwo}
\end{figure}

%Figures of beam contaminants in our beam PID line. 
%Table of beam purity reference to Jon's thesis paper
The secondary beam is produced through the projectile fragmentation of the primary beam off of a \SI{3}{\milli\metre} thick, rotating Be target \cite{inflightsep}. The resulting fragments are filtered in-flight to the desired seconary beam. The in-flight separation is handled by the BigRIPS fragment separator which is shown in Figure~\ref{fig:samuraiBeamLine}. The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$. Several sets of slits further purify the secondary beam quality by discarding particles which do not focus on the right focal planes. These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.  Each beam is tracked with the remaining part of the BigRIPS spectrometer tracking system. The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}. From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussian in this space.

@@

\begin{equation}
  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
   \label{multiGauss}
\end{equation}
where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{cccccc}
      \hline 
      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
      $\sigma_\mathrm{Z}$ & $\rho$\\
      \hline\hline 
      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian cuts parameters for all four systems beam PID.
    \label{beamParameters}}
  \end{center}
\end{table}

Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Figure~\ref{fig:beampid108} the ${}^{108}$Sn beam. The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events. For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.

@@

Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Figure~\ref{fig:edge}.  This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge. In this case the cluster position is biased towards the inside of the TPC. This also occurs for the first and last time bucket in the vertical direction. While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction. Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC. The hits that were cut out satisfied the following conditions,

@@

\begin{table}[!htb]
  \begin{center}
    \begin{tabular}{ccccc}
      \hline 
      System & $\mu_{\theta_\mathrm{a,proj}}$ &
      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
      $\sigma_{\theta_\mathrm{b,proj}}$ \\
      \hline\hline 
      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
      \hline
    \end{tabular}
    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
      $^{108}$Sn beam angle cuts. \label{beamAngleParameters}}
  \end{center}
\end{table}

@@

 The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame. The z-component of all runs in each beam type are plotted around the the target region in Figure~\ref{fig:vertexdist}. The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position. The mean position for each system is listed in Table~\ref{tb:vertexresol}. Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC. Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC. The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}. The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.

\begin{table*}\centering
\ra{1.3}
\begin{tabular}{@{}rrr@{}}\toprule
\multicolumn{3}{c}{Vertex Resolution}\\
\cmidrule{1-3}
System & Mean (cm) & Sigma (cm) \\
\midrule
$\tin{132}{124}$ & -14.79  & 1.2 \\
$\tin{124}{112}$ & -14.71  & 1.1 \\
$\tin{112}{124}$ & -14.78  & 1.2 \\
$\tin{108}{112}$ & -14.75  & 1.3 \\ 
\bottomrule
\end{tabular}
\caption{Summary of the mean vertex location for the target position in all systems and the measured vertex resolution.}
\label{tb:vertexresol}
\end{table*}

@@

\section{Impact Parameter Selection}
\label{sec:impact}

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=1.]{distancetovertex.pdf}
\label{fig:poca}
\caption{Example diagram of the Distance of Closest Approach (DOCA) cut for tracks.}
\end{figure}

@@

Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that effect the efficiency of track reconstruction. Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}. The relevant geometries which determine these regions are the corners of the field cage. Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given. Since the window to the field cage was slightly higher, the upper angles are smaller than the lower angles. These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.  While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Figure~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks. Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly. The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}. The values are very close to the simple calculation outlined previously.

@@

These regions can also be seen in the earlier discussion in the MC embedded efficiency in Figure~\ref{fig:pim_eff_ex}.  We could of course include a larger region where the efficiency is small, but not zero, but it is not good practice to  correct using small efficiency values. Instead we apply a reasonable $\phi$ cut in the good areas of reasonable efficiency and correct for the regions excluded from the cut. The solid angle covered by the cut is written as,
